\thetitle

### 2021-05-15 paper/paper.tex
rel lit

In contrast the standard discounted framework estimates only one single scalar value consisting of
the addition of these subterms, where however, the average reward is scaled by \(1/(1-\gamma)\).
%
To collect as much reward as possible, for instance to gain the highest expected profit, the
discount factor \(\gamma\) has to be very close to one, e.g. \(0.99\).
%
% As the discount factor \(\gamma\) is usually set very
% close to \(1\), e.g. \(0.99\), which
Such a parameterisation leads to the fact that the subterm including the average reward dominates
the other two.
%
Further, in commonly applied unichain (and thus ergodic) MDPs the average reward per step is equal
over all states. Thus, independently assessing it is not only computationally unwisely, but also
problematic when they are iteratively approximated. For this reasons most applications of discounted
reinforcement learning in literature evaluate their algorithms on processes that have an average
reward of about zero \citep{schneckenreither2020average}. \citet{schneckenreither2020average} shows
that tabular based discounted reinforcement learning agents are unable to find even basic policies
when confronted with processes that have a small average reward per step. Unfortunately, when
optimising for the highest profit (or the lowest cost), the most intuitive and clear way of reward
feedback to the agent is presenting it the profit that is accumulated every period.

\MS{todo}
%
Therefore, we adapt an algorithm that uses a scalar value for the estimation of the average reward
over all states, and estimates for every state-action pair that incorporate the bias value and the
sum, usually referred to as error term, of all other parts. The later values are adjusted by the
average reward.
% This, combined with the iterative evaluation and the independently assessing of the average reward
% for each state lets standard discounted reinforcement learning struggle to find good policies.
% Thus, this incautious combining of different kinds of state values as done in discounted
% reinforcement learning leads to the problems that (i) the average reward, which is equal for all
% states of usually investigated unichain MDPs, is dominating and thus diluting the bias values, and
% (ii) the state values are deteriorated by the error term that is only imposed due to the
% discounting technique.

\MS{why ARAL}

\MS{summary}

%
%

@@

An extreme scenario of this problem is the so called ``lead time syndrome'', which describes a
vicious cycle where increasing flow times perpetually inflate the lead times which leads to worse
performance (see e.g., \citealt{Mather1978, knollmann2013control, Selcuk2006}).
%
Thus, setting lead times dynamically harbors optimisation potential \citep{hoyt1978dynamic}, but may
also substantially degrade the system performance.
%
\citet{schneckenreither2020order} have established following categorisation of dynamic lead time
management approaches:

\begin{itemize}
\item \textsf{Reactive lead time management} approaches set lead times by reacting on
  earlier flow times (e.g., \citealt{enns2004work, Selcuk2006}). Note that the forecast is always
  based on \textit{past} data as the most recent system changes cannot be reflected by flow times
  until the corresponding orders arrive in the FGI, which might take several
  periods.
\item
  \textsf{Proactive lead time management} may incorporate \textit{past} data in conjunction with
  the \textit{current} system state to set lead times (e.g., \citealt{Bertrand1983,
    ChungHuang2002}). Put differently, these methods aim to find a function that provides lead times
  based on the current state of the production system and possibly information from the past.

\item \textsf{Predictive lead time management} may not only incorporate \textit{past} data and the
  \textit{current} system state to set lead times, but also utilises the anticipated \textit{future}
  system state to detect arising issues of future periods and react accordingly (e.g.
  \citealt{PaterninaArboleda2001,schneckenreither2020order,Schneckenreither2019}). Thus, it extends
  proactive lead time management from a flow time forecasting or simple lead time setting technique
  to a lead time management approach that integrates the future behaviour of the system when setting
  lead times. This allows reasoning of the system dynamics, as for instance triggering the lead time
  system, and thus such an algorithm can react accordingly to find a more farsighted optimal lead
  time update.
\end{itemize}

@@

Over the past decades reinforcement learning has been applied to various problems, for which
astonishing results have been reported. E.g.\@ only recently \citet{mnih2015human} have presented a
novel value-iteration reinforcement learning agent which exceeds human-level abilities in playing
many classic Atari 2600 games~\citep{bellemare2012investigating}. Further,
\citet{mnih2016asynchronous} present improved results with asynchronous actor-critic reinforcement
learning. Also games like Go~\citep{silver2016mastering} and Chess~\citep{silver2017mastering} have
been mastered with superhuman performance by \textit{tabula rasa} reinforcement learning agents.
Furthermore, the method has also been applied in the setting of manufacturing system, e.g.\@ to
improve the ramp-up process~\citep{doltsinis2012reinforcement}, in locally selecting appropriate
dispatching rules \citep{zhang1995reinforcement,wang2005application} or scheduling
\citep{zhang1995reinforcement, waschneck2018optimization}.
%
However, all these applications use discounted reinforcement learning and are either designed to
investigate a rather simple MDP, e.g.\@ by selecting heuristics instead of optimising the underlying
problem itself, or by mapping the actual objective in a reward function that is approximately \(0\)
on average over time.
%
This is due to the fact that state value is largely composed of a term defined by the policy's
average reward value \citep{MillerVeinott1969,Blackwell62} which would otherwise dilute the state
values and thus decrease the solution quality, as can for instance be observed in
\citet{SchneckenreitherHaeussler2019} and \citet{gijsbrechts2018can}.
%

@@

\MS[]{von hier}
Thus, as opposed to the commonly applied discounted reinforcement learning algorithm we use an
average reward adjusted reinforcement learning algorithm to adaptively release orders based on the
assessed state values of the production system. In contrast to the aforementioned works on average
reward reinforcement learning our algorithm incorporates the optimisation of not only the average
reward over time, but also the bias values, which is an important adaption in highly stochastic
systems. Only the work by
\citet{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}
integrates this second-level refinement optimisation. However, their algorithm requires the
selection of a reference state to prevent an infinite increase of state values. This results from
the lack of feedback in the iterative process of optimising the different refinement optimisation
levels. Furthermore, they asses the average reward independently for each state.

In summary, we propose a novel average reward adjusted reinforcement learning algorithm to assess
orders for their release to the shop floor. This is the first average reward adjusted reward
learning algorithm that operates using a artificial neural network as function approximation. To
ensure scalability we adapt a multi-agent approach, where each agent optimises the lead time
management of a single product type. The agents do so by assessing the expected costs for the
possible releases imposed by adapting the lead time. According to the learned estimates each agent
sets a planned lead time for the corresponding product type and with that releases orders into the
production system. The highly optimised value-iteration algorithm uses improved n-step reinforcement
learning with small action-based replay memories and worker agents to infer order release policies
which, especially in case of high utilisation, outperform not only the static order release
mechanisms, but also the commonly applied deep Q-Learning algorithm~\citep[cf.][]{mnih2015human}.

%
% \MS[t]{TODO} We provide ample evidence of the viability of the approach and compare it to well-known
% order release mechanisms. Furthermore, we show that the advantage of average reward reinforcement
% learning for the problem structures encountered in the context of production and logistics by
% comparing the results to the discounted reinforcement learning framework. The preliminary results
% show that our approach performs better than the other methods. With regard to practical implications
% we are confident that decision support tools based on average reward reinforcement learning increase
% the decision quality of human planners.

% \MS[t]{todo}
% Then~\ref{sec:Preliminary} presents preliminary
% result data, whereas Section~\ref{sec:conclusion} concludes this working paper.

>>>>

In contrast the standard discounted framework estimates only one single scalar value consisting of
the addition of these subterms, where however, the average reward is scaled by \(1/(1-\gamma)\).
%
To collect as much reward as possible, for instance to gain the highest expected profit, the
discount factor \(\gamma\) has to be very close to one, e.g. \(0.99\).
%
% As the discount factor \(\gamma\) is usually set very
% close to \(1\), e.g. \(0.99\), which
Such a parameterisation leads to the fact that the subterm including the average reward dominates
the other two.
%
Further, in commonly applied unichain (and thus ergodic) MDPs the average reward per step is equal
over all states. Thus, independently assessing it is not only computationally unwisely, but also
problematic when they are iteratively approximated. For this reasons most applications of discounted
reinforcement learning in literature evaluate their algorithms on processes that have an average
reward of about zero \citep{schneckenreither2020average}. \citet{schneckenreither2020average} shows
that tabular based discounted reinforcement learning agents are unable to find even basic policies
when confronted with processes that have a small average reward per step. Unfortunately, when
optimising for the highest profit (or the lowest cost) in the area of operations management, the
most intuitive and clear way of reward feedback to the agent is presenting it the profit that is
accumulated every period.

Therefore, we develop, adapt and enhance an algorithm that uses only one scalar value for the
estimation of the average reward over all states. Further it estimates for every state-action pair
the bias value and the sum of all other state value parts, usually referred to as error term. The
later values are adjusted by the average reward.
% This, combined with the iterative evaluation and the independently assessing of the average reward
% for each state lets standard discounted reinforcement learning struggle to find good policies.
% Thus, this incautious combining of different kinds of state values as done in discounted
% reinforcement learning leads to the problems that (i) the average reward, which is equal for all
% states of usually investigated unichain MDPs, is dominating and thus diluting the bias values, and
% (ii) the state values are deteriorated by the error term that is only imposed due to the
% discounting technique.
%
%
Thus, as opposed to the commonly applied discounted reinforcement learning algorithm we use an
average reward adjusted reinforcement learning algorithm to dynamically set lead time for order
release. In contrast to the aforementioned works on average reward reinforcement learning our
algorithm incorporates the optimisation of not only the average reward over time, but also the bias
values, which is an important adaption in highly stochastic systems. Only the work by
\citet{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}
integrates this second-level refinement optimisation. However, their average reward reinforcement
learning algorithm requires the selection of a reference state to prevent an infinite increase of
state values. This results from the lack of feedback in the iterative process of optimising
different sets of nested constraints. Furthermore, they asses the average reward independently for
each state.

To achieve a scalable lead time setting algorithm, we propose an extension that reduces the action
space tremendously. In fact, it transforms the problem size function measured on the number of
product types from one with an exponential increase to a linear one. The idea is that multiple
action decision agents are introduced that operate on the same state, but each one is responsible
for the lead time setting of only one product type.
% (i) To
% ensure scalability we adapt a multi-agent approach, where each agent optimises the lead time
% management of a single product type. The agents do so by assessing the expected costs for the
% possible releases imposed by adapting the lead time. According to the learned estimates each agent
% sets a planned lead time for the corresponding product type and with that releases orders into the
% production system.

Finally, we provide a proof-of-concept experimental section by testing the proposed algorithm on a
three stage flow shop with six product types, which provides ample evidence of the viability of the
approach.
% and are compared to static order release
% mechanisms. Furthermore, we show that the advantage of average reward reinforcement learning for the
% problem structures encountered in the context of production and logistics management by comparing
% the results to the discounted reinforcement learning
% framework.
The highly optimised value-iteration algorithm with improved n-step learning, small action-based
replay memories and concurrent workers infers order release policies which, especially in case of
high utilisation, outperform not only the static order release mechanisms, but also the commonly
applied deep Q-Learning algorithm~\citep[cf.][]{mnih2015human}.
% With regard to practical implications we are confident that decision support
% tools based on average reward reinforcement learning increase the decision quality of human
% planners.
%
In summary our contributions are as follows:
\begin{itemize}
\item We propose a novel \textit{average reward adjusted deep reinforcement learning algorithm} \ARA{}.
  This is the first average reward adjusted reward learning algorithm that operates using a
  artificial neural network as function approximation. It includes several known extensions like
  n-step learning \citep[e.g.][]{mnih2016asynchronous} or experience replay memory
  \citep[e.g.][]{mnih2015human}, which we advanced to prevent catastrophic forgetting by
  action-based selection of experiences.
\item We establish a action reduction method, named \textit{independent agents}, as a subset of the
  action space is treated as if there were multiple independent agents, although they share the same
  state. It reduces the growth of the number of possible actions from a exponential growth to a
  linear one. This is especially useful for product type based lead time management, as each agent
  is responsible for the lead time of one product type.
\item Using a three-stage flow shop model we show in an \textit{experimental evaluation} that the
  proposed algorithm outperforms standard Q-Learning and well established order release methods,
  especially for systems with high variability.
\end{itemize}

%
% \MS[t]{TODO}

% \MS[t]{todo}
% Then~\ref{sec:Preliminary} presents preliminary
% result data, whereas Section~\ref{sec:conclusion} concludes this working paper.
%
%

@@

Setting of lead times dynamically is a rather difficult task. Therefore, most state-of-the-art order
release mechanisms use static or fixed lead times to address the order release problem.
%
However, lead times should be set dynamically due to the non-linear relationship between utilisation
and flow times \citealt{Pahl2007,hoyt1978dynamic}, but this may also substantially degrade the
system performance. For instance, a solely reactive approach can result in the lead time syndrome.
%
The \textit{lead time syndrome} describes a vicious cycle where increasing flow times perpetually
inflate the lead times which leads to worse performance (see e.g., \citealt{Mather1978,
  knollmann2013control, Selcuk2006}).
%
%
\citet{schneckenreither2020order} have established following categorisation of dynamic lead time
management approaches:

\begin{itemize}
\item \textsf{Reactive lead time management} sets lead times by reacting on earlier flow times
  (e.g., \citealt{enns2004work, Selcuk2006}), i.e.\@ data from the past.
\item \textsf{Proactive lead time management} may incorporate \textit{past} data and the
  \textit{current} system state to set lead times (e.g., \citealt{Bertrand1983, ChungHuang2002}).
\item \textsf{Predictive lead time management} may incorporate \textit{past} data and the
  \textit{current} system state, and uses an anticipated \textit{future} system state to prevent
  arising issues of future periods (e.g.
  \citealt{PaterninaArboleda2001,schneckenreither2020order,Schneckenreither2019}).
\end{itemize}

@@

Over the past decades reinforcement learning has been applied to various problems, for which
astonishing results have been reported. E.g.\@ only recently \citet{mnih2015human} have presented a
novel value-iteration reinforcement learning agent which exceeds human-level abilities in playing
many classic Atari games~\citep{bellemare2012investigating}. Further, \citet{mnih2016asynchronous}
present improved results with asynchronous actor-critic reinforcement learning. Also games like
Go~\citep{silver2016mastering} and Chess~\citep{silver2017mastering} have been mastered with
superhuman performance by \textit{tabula rasa} reinforcement learning agents. Furthermore, the
method has also been applied in the setting of manufacturing system, e.g.\@ to improve the ramp-up
process~\citep{doltsinis2012reinforcement}, in locally selecting appropriate dispatching rules
\citep{zhang1995reinforcement,wang2005application} or scheduling \citep{zhang1995reinforcement,
  waschneck2018optimization}.
%
However, all these applications use discounted reinforcement learning and are either designed to
investigate a rather simple MDP, e.g.\@ by selecting heuristics instead of optimising the underlying
problem itself, or by mapping the actual objective in a reward function that is approximately \(0\)
on average over time.
%
This is due to the fact that state value is largely composed of a term defined by the policy's
average reward value \citep{MillerVeinott1969,Blackwell62} which would otherwise dilute the state
values and thus decrease the solution quality, as can for instance be observed in
\citet{SchneckenreitherHaeussler2019} and \citet{gijsbrechts2018can}.
%

@@

% Only the work by
% \citet{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}
% integrates this second-level refinement optimisation. However, their average reward reinforcement
% learning algorithm requires the selection of a reference state to prevent an infinite increase of
% state values. This results from the lack of feedback in the iterative process of optimising the
% different refinement optimisation levels. Furthermore, they asses the average reward independently
% for each state.
